---
title: "LF_08_clim_biome_ecoregion"
author: "Caroline McKeon"
date: "12/9/2020"
output: pdf_document
---

DATA NEEDED:


Data_08_anthromes_2_base_data_GeoTIFF/potveg.tif data from Ellis et al. (2010)
ecoregions based on Olson et al. (2001).


## set up
```{r}
library(tidyverse)
library(ggplot2)
library(raster)
library(rgdal)
library(data.table)
library(sjPlot) ## for the set_theme function
library(viridis)
library(ggpubr)
library(ggeffects)
library(gtools)
library(ggridges)
library(papeR)
library(kableExtra)
library(viridis)
library(gdalUtils)

## create logit transformation function
logitTransform <- function(x) { log(x/(1-x)) }
'%nin%' = Negate('%in%')
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message = FALSE, fig.width=16, fig.height=9.7) 
```

```{r}
# blank theme
set_theme(
  base = theme_classic(),
  axis.title.size = 2,
  axis.textsize = 1,
  legend.size = 1,
  legend.title.size = 1,
  geom.label.size = 1,
  plot.margin = margin(0.5,0.5, 0.5,0.5,"pt"))

## text editing
get_wraper <- function(width) {
  function(x) {
    lapply(strwrap(x, width = width, simplify = FALSE), paste, collapse="\n")}}
## colour palett
cb_pal <- c("#01665e", "#5ab4ac","#c7eae5", "#d8b365", 
            "#8c510a")
```

read in climate data
```{r}
mat <-raster('bio1.bil') ## mean annual temperature (C*10)
map <-raster('bio12.bil') ## mean annual precipatation (mm)
map_var <-raster('bio15.bil') ## mean annual precip coeff variation
mat_var <-raster('bio4.bil') ## mean annual temp SD*100
if(file.exists("Data_ModelDF_unscaled.rds")) {
    try(unscaled_data <- readRDS("Data_ModelDF_unscaled.rds"))
  } else warning("ModelDF_unscaled does not exist in this directory")
```


## Biomes
using data from Elis's anthromes to try figure out some average biome cliamte-clines
```{r eval=FALSE, include=FALSE}
biome_rast <- raster("Data_08_anthromes_2_base_data_GeoTIFF/potveg.tif")
biome_rast <- calc(biome_rast, fun=function(x){ x[x <= 0] <- NA; return(x)} )
plot(biome_rast, col=viridis_pal(option="C")(900))

biome_rast <- projectRaster(biome_rast, mat)
clim_map <- brick(map, mat, map_var, mat_var, biome_rast)

## #extract climate values for coordinates in (full) PREDICTS dataset
full_clim <- data.frame(raster::extract(clim_map, y=seq(1,7776000,1), df=TRUE)) 

full_clim <- unique(full_clim)
new <- full_clim[,2:6]
names(new) <- c("map", "mat", "map_var", "mat_var","biome") 
new <- new[new$biome >= 1,]
new <- drop_na(new)
new$biome <- round(new$biome, 0)

for (i in 1:12){
  new$map[new$biome == i] <- mean(new$map[new$biome == i])
  new$mat[new$biome == i] <- mean(new$mat[new$biome == i])
  new$map_var[new$biome == i] <- mean(new$map_var[new$biome == i])
  new$mat_var[new$biome == i] <- mean(new$mat_var[new$biome == i])
}

new <- unique(new)

b <- c("Tropical Evergreen Woodland", "Tropical Deciduous Woodland", "Temperate Evergreen Woodland", "Temperate Deciduous Woodland", 
"Boreal Woodland", "Mixed Woodland", "Savanna", "Grassland and Steppe", "Dense Shrubland", "Open Shrubland", "Tundra", "Deserts and Barren")

new <- new[order(new$biome),]
new$biome <- b
new$mat_var <- new$mat_var/100
new$mat <- new$mat/10
for (i in names(Filter(is.numeric, new))) {
new[,i] <- round(new[,i], 2)}
new <- new[order(-new$mat),]

biome$biome <- factor(biome$biome, 
                      levels = c("Tropical Evergreen Woodland", "Tropical Deciduous Woodland", 
                                 "Dense Shrubland", "Savanna", "Open Shrubland", "Deserts and Barren", 
                                 "Temperate Evergreen Woodland", "Temperate Deciduous Woodland", 
                                 "Grassland and Steppe", "Mixed Woodland", "Boreal Woodland", 
                                 "Tundra"))

#saveRDS(new, "Data_08_average_biome_clim_vars.rds")
```

```{r}
biome <- readRDS("Data_08_average_biome_clim_vars.rds")
```


## Ecoregions

```{r eval=FALSE, include=FALSE}
#eco <- shapefile("terr-ecoregions-TNC/tnc_terr_ecoregions.shp")

# teow <- readOGR(dsn = "Data_08_official", layer = "wwf_terr_ecos")
# # x <- rasterize(teow, mat)
# # writeRaster(x, "teow.grd")
# x <-gdalUtils::gdal_rasterize(teow, mat)
# levels(x)
# a <- deratify(x, "ECO_NAME")
#writeRaster(a, "Data_08_teow_econame.grd")
eco_rast <- raster("Data_08_teow_econame.grd")

clim_map <- brick(map, mat, map_var, mat_var, eco_rast)

## #extract climate values for coordinates in (full) PREDICTS dataset
full_clim <- data.frame(raster::extract(clim_map, y=seq(1,7776000,1), df=TRUE)) 

full_clim <- full_clim[which(!is.na(full_clim$ECO_NAME)),]
new <- full_clim[,-1]
new <- unique(new)
names(new) <- c("map", "mat", "map_var", "mat_var","eco") 
new <- drop_na(new)

#plot(eco_rast, col=viridis_pal(option="C")(900))

for (i in 1:826){
  print(i)
  new$map[new$eco == i] <- mean(new$map[new$eco == i])
  new$mat[new$eco == i] <- mean(new$mat[new$eco == i])/10
  new$map_var[new$eco == i] <- mean(new$map_var[new$eco == i])
  new$mat_var[new$eco == i] <- mean(new$mat_var[new$eco == i])/100
}

new <- unique(new)
new <- new[order(-new$mat),]
## drops 17 ecoregions (too coastal I assume)

#saveRDS(new, "Data_08_average_ecoregion_clim_vars.rds")
```


```{r}
eco <- readRDS("Data_08_average_ecoregion_clim_vars.rds")
```

## Average climatic differences between biomes and ecoregions
```{r}
average_diff <- function(clim_var, name, len){
  x <- data.frame()
  x <- as.data.frame(outer(clim_var,clim_var, "-"))
  names(x) <- levels(factor(name))
  rownames(x) <- levels(factor(name))
  
  ## get only lower triangle to avoid repetition
  x[lower.tri(x,diag=TRUE)] <- NA
  x <- x[,-1]
  x <- x[-len,]

  c <- c()
  n <- c()
  s <- as.data.frame(rep(rownames(x), (len-1)))
  for(i in names(x)){
    c <- append(c, x[,i])
    n <- append(n, rep(i, (len-1)))
    }
  s$diff <- c
  s$comp <- n
  s <- drop_na(s)
  return(s)
  }
```

```{r}
b <- as.data.frame(1:66)
vars <- c("map", "mat", "map_var", "mat_var")
for(j in names(biome[,which(names(biome) %in% vars)])){
  #print(j)
  b <- cbind(b, average_diff(biome[,j], biome$biome, 12))
}
names(b) <- c("1:66", "biome", "map", "comp", "rep(rownames(x), (len - 1))", 
              "mat", "comp", "rep(rownames(x), (len - 1))", "map_var", "comp", 
              "rep(rownames(x), (len - 1))", "mat_var", "comp_biome")
biome_diff <- b[,which(names(b) %in% c("biome", "map", "mat", "map_var", "mat_var", "comp_biome"))]
rm(b)
```

```{r}
e <- data.frame(1:326836)
vars <- c("map", "mat", "map_var", "mat_var")
for(j in names(eco[,which(names(eco) %in% vars)])){
  #print(j)
  e <- cbind(e, average_diff(eco[,j], eco$eco, 809))
}
names(e) <- c("X1.326836", "eco", "map", "comp", 
              "rep(rownames(x), (len - 1))", "mat", "comp", "rep(rownames(x), (len - 1))", 
              "map_var", "comp", "rep(rownames(x), (len - 1))", "mat_var", "comp_eco")
eco_diff <- e[,which(names(e) %in% c("eco","map", "mat", "map_var", "mat_var", "comp_eco"))]
rm(e)
```

```{r}
for (i in names(Filter(is.numeric, biome))) {
  biome[,i] <- round(biome[,i], 2)}
for (i in names(Filter(is.numeric, eco))) {
  eco[,i] <- round(eco[,i], 2)}

for (i in names(Filter(is.numeric, biome_diff))) {
  biome_diff[,i] <- round(biome_diff[,i], 2)}
for (i in names(Filter(is.numeric, eco_diff))) {
  eco_diff[,i] <- round(eco_diff[,i], 2)}
```

get modulus of differences
```{r}
for (i in names(Filter(is.numeric, biome_diff))) {
  biome_diff[,i] <- sqrt(biome_diff[,i]*biome_diff[,i])}
for (i in names(Filter(is.numeric, eco_diff))) {
  eco_diff[,i] <- sqrt(eco_diff[,i]*eco_diff[,i])}
```

get climate values used in model
```{r}
unscaled_vars <- c("unscaled_map", "unscaled_mat", "unscaled_map_var", "unscaled_mat_var")
data<- unscaled_data[,which(names(unscaled_data) %in% 
                              c("unscaled_map", "unscaled_mat", "unscaled_map_var", "unscaled_mat_var"))]
data<- data[,c("unscaled_map", "unscaled_mat", "unscaled_map_var", "unscaled_mat_var")]
data$unscaled_mat <- data$unscaled_mat/10
data$unscaled_mat_var <- data$unscaled_mat_var/100
names(data) <- vars
```

```{r}
# saveRDS(data, "Data_08_sd_model_clim_data.rds")
# saveRDS(biome_diff, "Data_08_average_biome_diffs.rds")
# saveRDS(eco_diff, "Data_08_average_ecoregion_diffs.rds")
```

```{r}
for(i in vars){
  print(i)
  print(length(eco_diff[,i][eco_diff[,i] <= sd(data[,i])])/length(eco_diff[,i]))
  }
```
## Plot
```{r eval=FALSE, include=FALSE}
par(mfrow =c(3,4), mar=c(4,2,3,2))

for(i in names(data[,which(names(data) %in% vars)])){
  hist(data[,i], freq = F,
       breaks = 300,
       border = "grey",
       ylab = NULL,
       main = paste("Model climate data"),
       xlab = paste(i))
  abline(v=mean(data[,i]),col="red", lwd = 2)
  abline(v=median(data[,i]),lty=2,col="red", lwd = 2)
  abline(v = mean(data[,i])+sd(data[,i]), lty = 1, col = "blue", lwd = 2)
  abline(v = mean(data[,i])-sd(data[,i]), lty = 1, col = "blue", lwd = 2)}
  # text(x = xmax, y = ymax, paste("Mean =", round(mean(data[,i]), 2), 
  #                   "SD =",round(sd(data[,i]), 2),
  #                    "Median =", round(median(data[,i]), 2)))}

for(i in names(biome_diff[,which(names(biome_diff) %in% vars)])){
  hist(biome_diff[,i],
       breaks = 300, 
       border = "grey",
        ylab = NULL,
       main = paste("biome difference"),
       xlab = paste(i))
    abline(v=mean(biome_diff[,i]),col="red", lwd = 2)
  abline(v=median(biome_diff[,i]),lty=2,col="red", lwd = 2)
  abline(v = sd(data[,i]), lty = 1, col = "blue", lwd = 2)}
# , "Mean =", round(mean(biome_diff[,i]), 2),
#                     "SD =",round(sd(biome_diff[,i]), 2),
#                      "Median =", round(median(biome_diff[,i]), 2)

for(i in names(eco_diff[,which(names(eco_diff) %in% vars)])){
  hist(eco_diff[,i],
       breaks = 300,
       border = "grey",
        ylab = NULL,
       main = paste("ecoregion difference"),
       xlab = paste(i))
    abline(v=mean(eco_diff[,i]),col="red", lwd = 2)
  abline(v=median(eco_diff[,i]),lty=2,col="red", lwd = 2)
  abline(v = sd(data[,i]), lty = 1, col = "blue", lwd = 2)}
# , "Mean =", round(mean(eco_diff[,i]), 2),
#                     "SD =",round(sd(eco_diff[,i]), 2),
#                      "Median =", round(median(eco_diff[,i]), 2)

```




```{r eval=FALSE, include=FALSE}
d <- data.frame(1)
for(i in names(data[,which(names(data) %in% vars)])){
 d[1,i] <- sd(data[,i])
 d[2,i] <- mean(data[,i])
 d[3,i] <- median(data[,i])}
d <- d[,-1]
names(d) <- vars
d <- rbind(c(rep(1,4)), d)
#rownames(d) <- c("data", "sd", "mean", "median")

b <- data.frame(1)
for(i in names(biome_diff[,which(names(biome_diff) %in% vars)])){
 b[1,i] <- sd(biome_diff[,i])
 b[2,i] <- mean(biome_diff[,i])
 b[3,i] <- median(biome_diff[,i])}
b<- b[,-1]
names(b) <- vars
b <- rbind(c(rep(1,4)), b)
#rownames(b) <- c("biome", "sd", "mean", "median")

e <- data.frame(1)
for(i in names(eco_diff[,which(names(eco_diff) %in% vars)])){
 e[1,i] <- sd(eco_diff[,i])
 e[2,i] <- mean(eco_diff[,i])
 e[3,i] <- median(eco_diff[,i])}
e<- e[,-1]
names(e) <- vars
e <- rbind(c(rep(1,4)), e)
#rownames(e) <- c("ecoregion", "sd", "mean", "median")

sum <- rbind(d,b,e)
for (i in names(Filter(is.numeric, sum))) {
  sum[,i] <- round(sum[,i], 2)}
sum$rn <-c("Model data", "sd", "mean", "median","Biome", "sd", "mean", "median", "Ecoregion", "sd", "mean", "median")
all_clim_ave <- sum[c(5,1:4)]
#saveRDS(all_clim_ave, "Data_08_all_clim_ave.rds")
```


