---
title: "LF_07b_NO_RICH_effectsizes"
author: "Caroline McKeon"
date: "2/27/2021"
output: word_document
---


## DATA NEEDED:

# Data_ModelDF.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R
# Data_ModelDF_unscaled.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R

## the three NO RICHNESS weighted effects coding models for the occurrence data, created in LF_04b_MO_RICH_frequentist_occurrence.R
# oc_wec_NO_RICH_int_maximal_zi_1_nested_no_P_C.rds
# oc_wec_NO_RICH_int_maximal_zi_1_nested_no_PF_P_C.rds
# oc_wec_NO_RICH_int_maximal_zi_1_nested_no_U_T.rds

## set up 
```{r set up, echo=FALSE}
library(DHARMa)
library(glmmTMB)
library(lme4)
library(data.table)
library(ggplot2)
library(sjPlot) ## for the set_theme function
library(viridis)
library(ggpubr)
library(ggeffects)
library(gtools)
library(ggridges)
library(papeR)
library(kableExtra)
library(tidyverse)

## create logit transformation function
logitTransform <- function(x) { log(x/(1-x)) }
'%nin%' = Negate('%in%')
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message = FALSE) 
knitr::opts_chunk$set(warning=FALSE) 
knitr::opts_chunk$set(fig.width=16, fig.height=9.7) 
```

# Occurrence
```{r}
## dummy coding version
  # if(file.exists("f_oc_maximal_zi_1_nested.rds")) {
  #   try(f_mod <- readRDS("f_oc_maximal_zi_1_nested.rdss"))
  # } else warning("f_oc_maximal_zi_1_nested.rds does not exist in this directory") ## 02/07/2020 all data

if(!exists("ModelDF")) {
  if(file.exists("Data_ModelDF.rds")) {
    try(ModelDF <- readRDS("Data_ModelDF.rds"))
  } else warning("ModelDF does not exist in this directory")
}

if(file.exists("Data_ModelDF_unscaled.rds")) {
    try(ModelDF_unscaled <- readRDS("Data_ModelDF_unscaled.rds"))
  } else warning("ModelDF_unscaled does not exist in this directory")

mydata <- ModelDF
## handle model dataframe to get species levels in the right format
levels(ModelDF$Best_guess_binomial) <- gsub(" ", "_", levels(ModelDF$Best_guess_binomial))
mydata$animal <- mydata$Best_guess_binomial

PR_oc <- readRDS("Data_03b_PR_f_oc.rds")
mydata <- droplevels(merge(mydata, PR_oc, by = "Best_guess_binomial",all.x = TRUE)) 

unscaled_data <- ModelDF_unscaled

levels(ModelDF_unscaled$Best_guess_binomial) <- gsub(" ", "_", levels(ModelDF$Best_guess_binomial))
unscaled_data$animal <- unscaled_data$Best_guess_binomial
unscaled_data <- droplevels(merge(unscaled_data, PR_oc, by = "Best_guess_binomial",all.x = TRUE)) 
```


```{r eval=FALSE, include=FALSE}
mean(mydata$pres_abs)
logitTransform(mean(mydata$pres_abs))
for (i in levels(mydata$Predominant_habitat)){
  print(c(i, length(mydata$Predominant_habitat[mydata$Predominant_habitat == i])))
}

for (i in levels(mydata$raunk_lf)){
  print(c(i, length(mydata$raunk_lf[mydata$raunk_lf == i])))
}
```

## Effect Size

```{r tidy oc wec a, eval=FALSE, include=FALSE}
f_mod_a <- readRDS("oc_wec_NO_RICH_int_maximal_zi_1_nested_no_U_T.rds")
f_oc_sum_a <- summary(f_mod_a)[["coefficients"]]
f_oc_sum_a <- f_oc_sum_a[c("cond","zi")]
f_oc_sum_a <- as.data.frame(f_oc_sum_a[["cond"]])
f_oc_sum_a <- setDT(f_oc_sum_a, keep.rownames = TRUE)[]

## adding a new column f_oc_sum_a that codes for if a term is landuse, lifeform or other
f_oc_sum_a$terms <- as.character(0)

## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_oc_sum_a$terms[grep(paste(landuse, collapse="|"), f_oc_sum_a$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_oc_sum_a$terms[grep(paste(clim, collapse="|"), f_oc_sum_a$rn)] <- "climate" 

## transforamation where all life forms are compared to relevant reference level 
f_oc_sum_a$eff_sz <- f_oc_sum_a$Estimate
f_oc_sum_a$eff_sz[2:14] <- f_oc_sum_a$eff_sz[2:14]+f_oc_sum_a$Estimate[1]
f_oc_sum_a$eff_sz[15:23] <- f_oc_sum_a$eff_sz[15:23]+f_oc_sum_a$eff_sz[11]+f_oc_sum_a$Estimate[2:10]
f_oc_sum_a$eff_sz[24:32] <- f_oc_sum_a$eff_sz[24:32]+f_oc_sum_a$eff_sz[12]+f_oc_sum_a$Estimate[2:10]
f_oc_sum_a$eff_sz[33:41] <- f_oc_sum_a$eff_sz[33:41]+f_oc_sum_a$eff_sz[13]+f_oc_sum_a$Estimate[2:10]
f_oc_sum_a$eff_sz[42:50] <- f_oc_sum_a$eff_sz[42:50]+f_oc_sum_a$eff_sz[14]+f_oc_sum_a$Estimate[2:10]
f_oc_sum_a$eff_sz[55:58] <- f_oc_sum_a$eff_sz[55:58]+f_oc_sum_a$eff_sz[51]
f_oc_sum_a$eff_sz[59:62] <- f_oc_sum_a$eff_sz[59:62]+f_oc_sum_a$eff_sz[52]
f_oc_sum_a$eff_sz[63:66] <- f_oc_sum_a$eff_sz[63:66]+f_oc_sum_a$eff_sz[53]
f_oc_sum_a$eff_sz[67:70] <- f_oc_sum_a$eff_sz[67:70]+f_oc_sum_a$eff_sz[54]


names(f_oc_sum_a)[names(f_oc_sum_a) == 'Std. Error'] <- 'Std.Error'
f_oc_sum_a$ci <- f_oc_sum_a$Std.Error*qnorm(.95)
f_oc_sum_a$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_oc_sum_a$landuse[grep(i, f_oc_sum_a$rn)] <- i }
  f_oc_sum_a$landuse[grep("indeterminate", f_oc_sum_a$rn)] <- "Secondary vegetation (indeterminate age)"

f_oc_sum_a$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_oc_sum_a$raunk_lf[grep(i, f_oc_sum_a$rn)] <- i }
  f_oc_sum_a$raunk_lf[grep("hemicryptophyte", f_oc_sum_a$rn)] <- "hemicryptophyte"

f_oc_sum_a$clim <- 0
covars <- c("humanfootprint_value", "map", "map_var", "mat",
            "mat_var")
for (i in covars){
  f_oc_sum_a$clim[grep(i, f_oc_sum_a$rn)] <- i }
  f_oc_sum_a$clim[grep("map_var", f_oc_sum_a$rn)] <- "map_var"
  f_oc_sum_a$clim[grep("mat_var", f_oc_sum_a$rn)] <- "mat_var"

f_oc_sum_a$raunk_lf <- factor(f_oc_sum_a$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_oc_sum_a$landuse <- factor(f_oc_sum_a$landuse, 
                      levels = c("Primary forest", "Primary non-forest", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Mature secondary vegetation", "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", 
                                 "Cropland", "Urban"))
f_oc_sum_a <- f_oc_sum_a[order(-f_oc_sum_a$raunk_lf, f_oc_sum_a$landuse),]
```

```{r tidy oc wec b, eval=FALSE, include=FALSE}
f_mod_b <- readRDS("oc_wec_NO_RICH_int_maximal_zi_1_nested_no_PF_P.rds")
f_oc_sum_b <- summary(f_mod_b)[["coefficients"]]
f_oc_sum_b <- f_oc_sum_b[c("cond","zi")]
f_oc_sum_b <- as.data.frame(f_oc_sum_b[["cond"]])
f_oc_sum_b <- setDT(f_oc_sum_b, keep.rownames = TRUE)[]

## adding a new column f_oc_sum_b that codes for if a term is landuse, lifeform or other
f_oc_sum_b$terms <- as.character(0)

## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_oc_sum_b$terms[grep(paste(landuse, collapse="|"), f_oc_sum_b$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_oc_sum_b$terms[grep(paste(clim, collapse="|"), f_oc_sum_b$rn)] <- "climate" 

## transforamation where all life forms are compared to relevant reference level 
f_oc_sum_b$eff_sz <- f_oc_sum_b$Estimate
f_oc_sum_b$eff_sz[2:14] <- f_oc_sum_b$eff_sz[2:14]+f_oc_sum_b$Estimate[1]
f_oc_sum_b$eff_sz[15:23] <- f_oc_sum_b$eff_sz[15:23]+f_oc_sum_b$eff_sz[11]+f_oc_sum_b$Estimate[2:10]
f_oc_sum_b$eff_sz[24:32] <- f_oc_sum_b$eff_sz[24:32]+f_oc_sum_b$eff_sz[12]+f_oc_sum_b$Estimate[2:10]
f_oc_sum_b$eff_sz[33:41] <- f_oc_sum_b$eff_sz[33:41]+f_oc_sum_b$eff_sz[13]+f_oc_sum_b$Estimate[2:10]
f_oc_sum_b$eff_sz[42:50] <- f_oc_sum_b$eff_sz[42:50]+f_oc_sum_b$eff_sz[14]+f_oc_sum_b$Estimate[2:10]
f_oc_sum_b$eff_sz[55:58] <- f_oc_sum_b$eff_sz[55:58]+f_oc_sum_b$eff_sz[51]
f_oc_sum_b$eff_sz[59:62] <- f_oc_sum_b$eff_sz[59:62]+f_oc_sum_b$eff_sz[52]
f_oc_sum_b$eff_sz[63:66] <- f_oc_sum_b$eff_sz[63:66]+f_oc_sum_b$eff_sz[53]
f_oc_sum_b$eff_sz[67:70] <- f_oc_sum_b$eff_sz[67:70]+f_oc_sum_b$eff_sz[54]


names(f_oc_sum_b)[names(f_oc_sum_b) == 'Std. Error'] <- 'Std.Error'
f_oc_sum_b$ci <- f_oc_sum_b$Std.Error*qnorm(.95)
f_oc_sum_b$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_oc_sum_b$landuse[grep(i, f_oc_sum_b$rn)] <- i }
  f_oc_sum_b$landuse[grep("indeterminate", f_oc_sum_b$rn)] <- "Secondary vegetation (indeterminate age)"

f_oc_sum_b$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_oc_sum_b$raunk_lf[grep(i, f_oc_sum_b$rn)] <- i }
  f_oc_sum_b$raunk_lf[grep("hemicryptophyte", f_oc_sum_b$rn)] <- "hemicryptophyte"

f_oc_sum_b$clim <- 0
covars <- c("humanfootprint_value", "map", "map_var", "mat", 
            "mat_var")
for (i in covars){
  f_oc_sum_b$clim[grep(i, f_oc_sum_b$rn)] <- i }
  f_oc_sum_b$clim[grep("map_var", f_oc_sum_b$rn)] <- "map_var"
  f_oc_sum_b$clim[grep("mat_var", f_oc_sum_b$rn)] <- "mat_var"

f_oc_sum_b$raunk_lf <- factor(f_oc_sum_b$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_oc_sum_b$landuse <- factor(f_oc_sum_b$landuse, 
                      levels = c("Primary forest", "Primary non-forest", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Mature secondary vegetation", "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", 
                                 "Cropland", "Urban"))
f_oc_sum_b <- f_oc_sum_b[order(-f_oc_sum_b$raunk_lf, f_oc_sum_b$landuse),]
```

```{r tidy oc wec c, eval=FALSE, include=FALSE}
f_mod_c <- readRDS("oc_wec_NO_RICH_int_maximal_zi_1_nested_no_P_C.rds")
f_oc_sum_c <- summary(f_mod_c)[["coefficients"]]
f_oc_sum_c <- f_oc_sum_c[c("cond","zi")]
f_oc_sum_c <- as.data.frame(f_oc_sum_c[["cond"]])
f_oc_sum_c <- setDT(f_oc_sum_c, keep.rownames = TRUE)[]

## adding a new column f_oc_sum_c that codes for if a term is landuse, lifeform or other
f_oc_sum_c$terms <- as.character(0)

## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_oc_sum_c$terms[grep(paste(landuse, collapse="|"), f_oc_sum_c$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_oc_sum_c$terms[grep(paste(clim, collapse="|"), f_oc_sum_c$rn)] <- "climate" 

## transforamation where all life forms are compared to relevant reference level 
f_oc_sum_c$eff_sz <- f_oc_sum_c$Estimate
f_oc_sum_c$eff_sz[2:14] <- f_oc_sum_c$eff_sz[2:14]+f_oc_sum_c$Estimate[1]
f_oc_sum_c$eff_sz[15:23] <- f_oc_sum_c$eff_sz[15:23]+f_oc_sum_c$eff_sz[11]+f_oc_sum_c$Estimate[2:10]
f_oc_sum_c$eff_sz[24:32] <- f_oc_sum_c$eff_sz[24:32]+f_oc_sum_c$eff_sz[12]+f_oc_sum_c$Estimate[2:10]
f_oc_sum_c$eff_sz[33:41] <- f_oc_sum_c$eff_sz[33:41]+f_oc_sum_c$eff_sz[13]+f_oc_sum_c$Estimate[2:10]
f_oc_sum_c$eff_sz[42:50] <- f_oc_sum_c$eff_sz[42:50]+f_oc_sum_c$eff_sz[14]+f_oc_sum_c$Estimate[2:10]
f_oc_sum_c$eff_sz[55:58] <- f_oc_sum_c$eff_sz[55:58]+f_oc_sum_c$eff_sz[51]
f_oc_sum_c$eff_sz[59:62] <- f_oc_sum_c$eff_sz[59:62]+f_oc_sum_c$eff_sz[52]
f_oc_sum_c$eff_sz[63:66] <- f_oc_sum_c$eff_sz[63:66]+f_oc_sum_c$eff_sz[53]
f_oc_sum_c$eff_sz[67:70] <- f_oc_sum_c$eff_sz[67:70]+f_oc_sum_c$eff_sz[54]


names(f_oc_sum_c)[names(f_oc_sum_c) == 'Std. Error'] <- 'Std.Error'
f_oc_sum_c$ci <- f_oc_sum_c$Std.Error*qnorm(.95)
f_oc_sum_c$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_oc_sum_c$landuse[grep(i, f_oc_sum_c$rn)] <- i }
  f_oc_sum_c$landuse[grep("indeterminate", f_oc_sum_c$rn)] <- "Secondary vegetation (indeterminate age)"

f_oc_sum_c$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_oc_sum_c$raunk_lf[grep(i, f_oc_sum_c$rn)] <- i }
  f_oc_sum_c$raunk_lf[grep("hemicryptophyte", f_oc_sum_c$rn)] <- "hemicryptophyte"

f_oc_sum_c$clim <- 0
covars <- c("humanfootprint_value", "map", "map_var", "mat", 
            "mat_var")
for (i in covars){
  f_oc_sum_c$clim[grep(i, f_oc_sum_c$rn)] <- i }
  f_oc_sum_c$clim[grep("map_var", f_oc_sum_c$rn)] <- "map_var"
  f_oc_sum_c$clim[grep("mat_var", f_oc_sum_c$rn)] <- "mat_var"

f_oc_sum_c$raunk_lf <- factor(f_oc_sum_c$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_oc_sum_c$landuse <- factor(f_oc_sum_c$landuse, 
                      levels = c("Primary forest", "Primary non-forest", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Mature secondary vegetation", "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", 
                                 "Cropland", "Urban"))
f_oc_sum_c <- f_oc_sum_c[order(-f_oc_sum_c$raunk_lf, f_oc_sum_c$landuse),]
```

```{r, eval=FALSE, include=FALSE}
f_oc_sum <- unique(rbind(f_oc_sum_a, f_oc_sum_b, f_oc_sum_c))
f_oc_sum$Estimate <- round(f_oc_sum$Estimate, 3)
f_oc_sum$Std.Error <- round(f_oc_sum$Std.Error, 3)
f_oc_sum$`z value` <- round(f_oc_sum$`z value`, 3)
f_oc_sum$`Pr(>|z|)` <- round(f_oc_sum$`Pr(>|z|)`, 3)
f_oc_sum$eff_sz <- round(f_oc_sum$eff_sz, 3)
f_oc_sum$ci <- round(f_oc_sum$ci, 3)
f_oc_sum <- unique(f_oc_sum)
```


```{r}
#get average of estimates that differ at the 3rd decimal place
for (i in levels(factor(f_oc_sum$rn))){
  if (length(f_oc_sum$rn[f_oc_sum$rn == i]) > 1) {
  f_oc_sum$Estimate[f_oc_sum$rn == i] <- mean(f_oc_sum$Estimate[f_oc_sum$rn == i])
  f_oc_sum$Std.Error[f_oc_sum$rn == i] <- mean(f_oc_sum$Std.Error[f_oc_sum$rn == i])
  f_oc_sum$`z value`[f_oc_sum$rn == i] <- mean(f_oc_sum$`z value`[f_oc_sum$rn == i])
  f_oc_sum$`Pr(>|z|)`[f_oc_sum$rn == i] <- mean(f_oc_sum$`Pr(>|z|)`[f_oc_sum$rn == i])
  f_oc_sum$eff_sz[f_oc_sum$rn == i] <- mean(f_oc_sum$eff_sz[f_oc_sum$rn == i])
  f_oc_sum$ci[f_oc_sum$rn == i] <- mean(f_oc_sum$ci[f_oc_sum$rn == i])
  }
}
f_oc_sum <- unique(f_oc_sum)
```

make the column for the histogram
```{r, eval=FALSE, include=FALSE}
for(i in levels(f_oc_sum$landuse)){
f_oc_sum$luN[f_oc_sum$landuse == i] <- 
  length(mydata$Predominant_habitat[mydata$Predominant_habitat == i])}
```

quick cheat for knitting plots
```{r}
#saveRDS(f_oc_sum, "f_oc_NO_RICH_sum.rds")
f_oc_sum <- readRDS("f_oc_NO_RICH_sum.rds")

f_oc_sum$land_av <- NA
f_oc_sum$land_av[which(!is.na(f_oc_sum$landuse) & is.na(f_oc_sum$raunk_lf))] <- 1
#saveRDS(f_oc_sum, "f_oc_NO_RICH_sum.rds")
```




