---
title: "LF_13_merged_pc_estimates"
author: "Caroline McKeon"
date: '2023-01-12'
output: word_document
---

## DATA NEEDED:

# Data_ModelDF.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R
# Data_ModelDF_unscaled.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R

## the three weighted effects coding models for the percent cover data, created in LF_12_merged_LU_frequentist_percent_cover.R
# pc_merged_no_P_C.rds
# pc_merged_no_PV_P.rds
# pc_merged_no_U_T.rds



## set up 
```{r set up, echo=FALSE}
library(glmmTMB)
library(lme4)
library(data.table)
library(ggplot2)
library(sjPlot) ## for the set_theme function
library(viridis)
library(ggpubr)
library(ggeffects)
library(gtools)
library(ggridges)
library(papeR)
library(kableExtra)

## create logit transformation function
logitTransform <- function(x) { log(x/(1-x)) }
'%nin%' = Negate('%in%')
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message = FALSE) 
knitr::opts_chunk$set(warning=FALSE) 
knitr::opts_chunk$set(fig.width=14, fig.height=7) 
```

```{r read in models}

if(!exists("ModelDF")) {
  if(file.exists("Data_ModelDF.rds")) {
    try(ModelDF <- readRDS("Data_ModelDF.rds"))
  } else warning("ModelDF does not exist in this directory")
}

if(file.exists("Data_ModelDF_unscaled.rds")) {
    try(ModelDF_unscaled <- readRDS("Data_ModelDF_unscaled.rds"))
  } else warning("ModelDF_unscaled does not exist in this directory")
```

```{r handle model dataframe}
mydata <- ModelDF

## handle model dataframe to get just percent cover data, with species levels in the right format
levels(ModelDF$Best_guess_binomial) <- gsub(" ", "_", levels(ModelDF$Best_guess_binomial))
mydata <- ModelDF[ModelDF$Diversity_metric == "percent cover", ]
mydata <- mydata[mydata$Measurement !=0,]
mydata <- unique(mydata)
mydata$Measurement <- mydata$Measurement/100
mydata$response <- scale(logitTransform(mydata$Measurement))
mydata$animal <- mydata$Best_guess_binomial

PR_pc <- readRDS("Data_03a_PR_f_pc.rds")
mydata <- droplevels(merge(mydata, PR_pc, by = "Best_guess_binomial",all.x = TRUE)) 

unscaled_data <- ModelDF_unscaled

levels(ModelDF_unscaled$Best_guess_binomial) <- gsub(" ", "_", levels(ModelDF$Best_guess_binomial))
unscaled_data <- ModelDF_unscaled[ModelDF_unscaled$Diversity_metric == "percent cover", ]
unscaled_data <- unscaled_data[unscaled_data$Measurement !=0,]
unscaled_data <- unique(unscaled_data)
unscaled_data$Measurement <- unscaled_data$Measurement/100
unscaled_data$response <- scale(logitTransform(unscaled_data$Measurement))
unscaled_data$animal <- unscaled_data$Best_guess_binomial

unscaled_data <- droplevels(merge(unscaled_data, PR_pc, by = "Best_guess_binomial",all.x = TRUE)) 

## merge primary land uses

mydata$Predominant_habitat <- as.character(mydata$Predominant_habitat)
mydata$Predominant_habitat[mydata$Predominant_habitat %in% c("Primary forest", "Primary non-forest")] <- "Primary vegetation"
mydata$Predominant_habitat <- factor(mydata$Predominant_habitat)
```


```{r}
## read in models
f_mod_a <- readRDS("pc_merged_no_P_C.rds")
f_mod_b <- readRDS("pc_merged_no_PV_P.rds")
f_mod_c <- readRDS("pc_merged_no_U_T.rds")

```


## Model Estimates

```{r wec int a handle labling and tidy format pc}
f_pc_sum_a <- as.data.frame(summary(f_mod_a)[["coefficients"]])
f_pc_sum_a <- setDT(f_pc_sum_a, keep.rownames = TRUE)[]

## adding a new column f_pc_sum_a that codes for if a term is landuse, lifeform or other
f_pc_sum_a$terms <- as.character(0)

## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_pc_sum_a$terms[grep(paste(landuse, collapse = "|"), f_pc_sum_a$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_pc_sum_a$terms[grep(paste(clim, collapse="|"), f_pc_sum_a$rn)] <- "climate" 
spp <- c("Species_richness", "spp")
f_pc_sum_a$terms[grep(paste(spp, collapse = "|"), f_pc_sum_a$rn)] <- "species richness"

## transforamation where all life forms are compared to relevant reference level 
## wec_int calculation
f_pc_sum_a$eff_sz <- f_pc_sum_a$Estimate
f_pc_sum_a$eff_sz[2:12] <- f_pc_sum_a$eff_sz[2:12]+f_pc_sum_a$Estimate[1]
f_pc_sum_a$eff_sz[13:19] <- f_pc_sum_a$eff_sz[13:19]+f_pc_sum_a$eff_sz[9]+f_pc_sum_a$Estimate[2:8]
f_pc_sum_a$eff_sz[20:26] <- f_pc_sum_a$eff_sz[20:26]+f_pc_sum_a$eff_sz[10]+f_pc_sum_a$Estimate[2:8]
f_pc_sum_a$eff_sz[27:33] <- f_pc_sum_a$eff_sz[27:33]+f_pc_sum_a$eff_sz[11]+f_pc_sum_a$Estimate[2:8]
f_pc_sum_a$eff_sz[34:40] <- f_pc_sum_a$eff_sz[34:40]+f_pc_sum_a$eff_sz[12]+f_pc_sum_a$Estimate[2:8]
f_pc_sum_a$eff_sz[46:49] <- f_pc_sum_a$eff_sz[46:49]+f_pc_sum_a$eff_sz[41]
f_pc_sum_a$eff_sz[50:53] <- f_pc_sum_a$eff_sz[50:53]+f_pc_sum_a$eff_sz[42]
f_pc_sum_a$eff_sz[54:57] <- f_pc_sum_a$eff_sz[54:57]+f_pc_sum_a$eff_sz[43]
f_pc_sum_a$eff_sz[58:61] <- f_pc_sum_a$eff_sz[58:61]+f_pc_sum_a$eff_sz[44]
f_pc_sum_a$eff_sz[62:65] <- f_pc_sum_a$eff_sz[62:65]+f_pc_sum_a$eff_sz[55]


names(f_pc_sum_a)[names(f_pc_sum_a) == 'Std. Error'] <- 'Std.Error'
f_pc_sum_a$ci <- f_pc_sum_a$`Std.Error`*qnorm(.95)
f_pc_sum_a$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_pc_sum_a$landuse[grep(i, f_pc_sum_a$rn)] <- i }
  f_pc_sum_a$landuse[grep("(indeterminate age)", f_pc_sum_a$rn)] <- "Secondary vegetation (indeterminate age)"

f_pc_sum_a$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_pc_sum_a$raunk_lf[grep(i, f_pc_sum_a$rn)] <- i }
  f_pc_sum_a$raunk_lf[grep("hemicryptophyte", f_pc_sum_a$rn)] <- "hemicryptophyte"

f_pc_sum_a$clim <- 0
covars <- c("map", "map_var", "Species_richness", "mat", "spp",
            "mat_var")
for (i in covars){
  f_pc_sum_a$clim[grep(i, f_pc_sum_a$rn)] <- i }
  f_pc_sum_a$clim[grep("map_var", f_pc_sum_a$rn)] <- "map_var"
  f_pc_sum_a$clim[grep("mat_var", f_pc_sum_a$rn)] <- "mat_var"

f_pc_sum_a$clim[f_pc_sum_a$terms== "species richness"] <- "species richness"

f_pc_sum_a$raunk_lf <- factor(f_pc_sum_a$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_pc_sum_a$landuse <- factor(f_pc_sum_a$landuse, 
                      levels = c("Primary vegetation", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", "Cropland", "Urban"))
f_pc_sum_a <- f_pc_sum_a[order(-f_pc_sum_a$raunk_lf, f_pc_sum_a$landuse),]

```

```{r wec int b handle labling and tidy format pc}
f_pc_sum_b <- as.data.frame(summary(f_mod_b)[["coefficients"]])
f_pc_sum_b <- setDT(f_pc_sum_b, keep.rownames = TRUE)[]

## adding a new column f_pc_sum_b that codes for if a term is landuse, lifeform or other
f_pc_sum_b$terms <- as.character(0)
## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_pc_sum_b$terms[grep(paste(landuse, collapse = "|"), f_pc_sum_b$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_pc_sum_b$terms[grep(paste(clim, collapse="|"), f_pc_sum_b$rn)] <- "climate" 
spp <- c("Species_richness", "spp")
f_pc_sum_b$terms[grep(paste(spp, collapse = "|"), f_pc_sum_b$rn)] <- "species richness"

## transforamation where all life forms are compared to relevant reference level 
## wec_int calculation
f_pc_sum_b$eff_sz <- f_pc_sum_b$Estimate
f_pc_sum_b$eff_sz[2:12] <- f_pc_sum_b$eff_sz[2:12]+f_pc_sum_b$Estimate[1]
f_pc_sum_b$eff_sz[13:19] <- f_pc_sum_b$eff_sz[13:19]+f_pc_sum_b$eff_sz[9]+f_pc_sum_b$Estimate[2:8]
f_pc_sum_b$eff_sz[20:26] <- f_pc_sum_b$eff_sz[20:26]+f_pc_sum_b$eff_sz[10]+f_pc_sum_b$Estimate[2:8]
f_pc_sum_b$eff_sz[27:33] <- f_pc_sum_b$eff_sz[27:33]+f_pc_sum_b$eff_sz[11]+f_pc_sum_b$Estimate[2:8]
f_pc_sum_b$eff_sz[34:40] <- f_pc_sum_b$eff_sz[34:40]+f_pc_sum_b$eff_sz[12]+f_pc_sum_b$Estimate[2:8]
f_pc_sum_b$eff_sz[46:49] <- f_pc_sum_b$eff_sz[46:49]+f_pc_sum_b$eff_sz[41]
f_pc_sum_b$eff_sz[50:53] <- f_pc_sum_b$eff_sz[50:53]+f_pc_sum_b$eff_sz[42]
f_pc_sum_b$eff_sz[54:57] <- f_pc_sum_b$eff_sz[54:57]+f_pc_sum_b$eff_sz[43]
f_pc_sum_b$eff_sz[58:61] <- f_pc_sum_b$eff_sz[58:61]+f_pc_sum_b$eff_sz[44]
f_pc_sum_b$eff_sz[62:65] <- f_pc_sum_b$eff_sz[62:65]+f_pc_sum_b$eff_sz[55]

names(f_pc_sum_b)[names(f_pc_sum_b) == 'Std. Error'] <- 'Std.Error'
f_pc_sum_b$ci <- f_pc_sum_b$`Std.Error`*qnorm(.95)
f_pc_sum_b$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_pc_sum_b$landuse[grep(i, f_pc_sum_b$rn)] <- i }
  f_pc_sum_b$landuse[grep("(indeterminate age)", f_pc_sum_b$rn)] <- "Secondary vegetation (indeterminate age)"

f_pc_sum_b$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_pc_sum_b$raunk_lf[grep(i, f_pc_sum_b$rn)] <- i }
  f_pc_sum_b$raunk_lf[grep("hemicryptophyte", f_pc_sum_b$rn)] <- "hemicryptophyte"

f_pc_sum_b$clim <- 0
covars <- c("map", "map_var", "Species_richness", "mat", "spp",
            "mat_var")
for (i in covars){
  f_pc_sum_b$clim[grep(i, f_pc_sum_b$rn)] <- i }
  f_pc_sum_b$clim[grep("map_var", f_pc_sum_b$rn)] <- "map_var"
  f_pc_sum_b$clim[grep("mat_var", f_pc_sum_b$rn)] <- "mat_var"

f_pc_sum_b$clim[f_pc_sum_b$terms== "species richness"] <- "species richness"

f_pc_sum_b$raunk_lf <- factor(f_pc_sum_b$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_pc_sum_b$landuse <- factor(f_pc_sum_b$landuse, 
                      levels = c("Primary vegetation", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", "Cropland", "Urban"))
f_pc_sum_b <- f_pc_sum_b[order(-f_pc_sum_b$raunk_lf, f_pc_sum_b$landuse),]

```

```{r wec int c handle labling and tidy format pc}
f_pc_sum_c <- as.data.frame(summary(f_mod_c)[["coefficients"]])
f_pc_sum_c <- setDT(f_pc_sum_c, keep.rownames = TRUE)[]

## adding a new column f_pc_sum_c that codes for if a term is landuse, lifeform or other
f_pc_sum_c$terms <- as.character(0)
## Handling long term-names
landuse <- c("Predominant_habitat.wec", "hab_raunk_interactionx1")
f_pc_sum_c$terms[grep(paste(landuse, collapse = "|"), f_pc_sum_c$rn)] <- "land use"
clim <- c("mat_var","map", "map_var", "mat")
f_pc_sum_c$terms[grep(paste(clim, collapse="|"), f_pc_sum_c$rn)] <- "climate" 
spp <- c("Species_richness", "spp")
f_pc_sum_c$terms[grep(paste(spp, collapse = "|"), f_pc_sum_c$rn)] <- "species richness"

## transforamation where all life forms are compared to relevant reference level 
## wec_int calculation
f_pc_sum_c$eff_sz <- f_pc_sum_c$Estimate
f_pc_sum_c$eff_sz[2:12] <- f_pc_sum_c$eff_sz[2:12]+f_pc_sum_c$Estimate[1]
f_pc_sum_c$eff_sz[13:19] <- f_pc_sum_c$eff_sz[13:19]+f_pc_sum_c$eff_sz[9]+f_pc_sum_c$Estimate[2:8]
f_pc_sum_c$eff_sz[20:26] <- f_pc_sum_c$eff_sz[20:26]+f_pc_sum_c$eff_sz[10]+f_pc_sum_c$Estimate[2:8]
f_pc_sum_c$eff_sz[27:33] <- f_pc_sum_c$eff_sz[27:33]+f_pc_sum_c$eff_sz[11]+f_pc_sum_c$Estimate[2:8]
f_pc_sum_c$eff_sz[34:40] <- f_pc_sum_c$eff_sz[34:40]+f_pc_sum_c$eff_sz[12]+f_pc_sum_c$Estimate[2:8]
f_pc_sum_c$eff_sz[46:49] <- f_pc_sum_c$eff_sz[46:49]+f_pc_sum_c$eff_sz[41]
f_pc_sum_c$eff_sz[50:53] <- f_pc_sum_c$eff_sz[50:53]+f_pc_sum_c$eff_sz[42]
f_pc_sum_c$eff_sz[54:57] <- f_pc_sum_c$eff_sz[54:57]+f_pc_sum_c$eff_sz[43]
f_pc_sum_c$eff_sz[58:61] <- f_pc_sum_c$eff_sz[58:61]+f_pc_sum_c$eff_sz[44]
f_pc_sum_c$eff_sz[62:65] <- f_pc_sum_c$eff_sz[62:65]+f_pc_sum_c$eff_sz[55]

names(f_pc_sum_c)[names(f_pc_sum_c) == 'Std. Error'] <- 'Std.Error'
f_pc_sum_c$ci <- f_pc_sum_c$`Std.Error`*qnorm(.95)
f_pc_sum_c$landuse <- 0
for (i in unique(levels(mydata$Predominant_habitat))){
  f_pc_sum_c$landuse[grep(i, f_pc_sum_c$rn)] <- i }
  f_pc_sum_c$landuse[grep("(indeterminate age)", f_pc_sum_c$rn)] <- "Secondary vegetation (indeterminate age)"

f_pc_sum_c$raunk_lf <- 0
for (i in unique(levels(mydata$raunk_lf))){
  f_pc_sum_c$raunk_lf[grep(i, f_pc_sum_c$rn)] <- i }
  f_pc_sum_c$raunk_lf[grep("hemicryptophyte", f_pc_sum_c$rn)] <- "hemicryptophyte"

f_pc_sum_c$clim <- 0
covars <- c("map", "map_var", "Species_richness", "mat", "spp",
            "mat_var")
for (i in covars){
  f_pc_sum_c$clim[grep(i, f_pc_sum_c$rn)] <- i }
  f_pc_sum_c$clim[grep("map_var", f_pc_sum_c$rn)] <- "map_var"
  f_pc_sum_c$clim[grep("mat_var", f_pc_sum_c$rn)] <- "mat_var"

f_pc_sum_c$clim[f_pc_sum_c$terms== "species richness"] <- "species richness"

f_pc_sum_c$raunk_lf <- factor(f_pc_sum_c$raunk_lf , 
                      levels = c("phanerophyte","chamaephyte", "hemicryptophyte", "cryptophyte", 
                                 "therophyte"))
f_pc_sum_c$landuse <- factor(f_pc_sum_c$landuse, 
                      levels = c("Primary vegetation", "Young secondary vegetation", "Intermediate secondary vegetation", 
                                 "Secondary vegetation (indeterminate age)", "Plantation forest", "Pasture", "Cropland", "Urban"))
f_pc_sum_c <- f_pc_sum_c[order(-f_pc_sum_c$raunk_lf, f_pc_sum_c$landuse),]

```

```{r}
f_pc_sum <- unique(rbind(f_pc_sum_a, f_pc_sum_b, f_pc_sum_c))
f_pc_sum$Estimate <- round(f_pc_sum$Estimate, 3)
f_pc_sum$Std.Error <- round(f_pc_sum$Std.Error, 3)
f_pc_sum$`t value` <- round(f_pc_sum$`t value`, 3)
f_pc_sum$eff_sz <- round(f_pc_sum$eff_sz, 3)
f_pc_sum$ci <- round(f_pc_sum$ci, 3)
f_pc_sum <- unique(f_pc_sum)
```

```{r}
#get average of estimates that differ at the 3rd decimal place
for (i in levels(factor(f_pc_sum$rn))){
  if (length(f_pc_sum$rn[f_pc_sum$rn == i]) > 1) {
  f_pc_sum$Estimate[f_pc_sum$rn == i] <- mean(f_pc_sum$Estimate[f_pc_sum$rn == i])
  f_pc_sum$Std.Error[f_pc_sum$rn == i] <- mean(f_pc_sum$Std.Error[f_pc_sum$rn == i])
  f_pc_sum$`t value`[f_pc_sum$rn == i] <- mean(f_pc_sum$`t value`[f_pc_sum$rn == i])
  f_pc_sum$eff_sz[f_pc_sum$rn == i] <- mean(f_pc_sum$eff_sz[f_pc_sum$rn == i])
  f_pc_sum$ci[f_pc_sum$rn == i] <- mean(f_pc_sum$ci[f_pc_sum$rn == i])
  }
}
f_pc_sum <- unique(f_pc_sum)
```


make the column for the histograms which will be needed for nice graphs
```{r}
for(i in levels(f_pc_sum$landuse)){
f_pc_sum$luN[f_pc_sum$landuse == i] <- 
  length(mydata$Predominant_habitat[mydata$Predominant_habitat == i])}

for(i in levels(f_pc_sum$raunk_lf)){
f_pc_sum$raunkN[f_pc_sum$raunk_lf == i] <- 
  length(mydata$raunk_lf[mydata$raunk_lf == i])}
```


```{r}

#f_pc_sum <- readRDS("f_pc_sum.rds")
f_pc_sum <- f_pc_sum[order(-f_pc_sum$raunk_lf, f_pc_sum$landuse),]
f_pc_sum$land_av <- 0
f_pc_sum$land_av[which(is.na(f_pc_sum$raunk_lf) & !is.na(f_pc_sum$landuse))] <- 1
#saveRDS(f_pc_sum, "f_pc_sum_merged.rds")
```

