---
title: "LF_10_sup_figures"
author: "Caroline McKeon"
date: "10/19/2021"
output: word_document
---

## DATA NEEDED:

# f_oc_sum - dataframe of results from occurrence model object - created in LF_06b_oc_estimates.R
# f_pc_sum - dataframe of results from abundance model object - created in LF_06a_pc_estimates.R

# Data_ModelDF.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R

# Data_08_anthromes_2_base_data_GeoTIFF/potveg.tif data from Ellis et al. (2010)
# Data_08_average_biome_clim_vars.rds created in LF_08_clim_biome_ecoregion.Rmd
# Data_08_average_ecoregion_clim_vars.rds created in LF_08_clim_biome_ecoregion.Rmd

# Data_08_sd_model_clim_data.rds created in LF_08_clim_biome_ecoregion.Rmd
# Data_08_average_biome_diffs.rds created in LF_08_clim_biome_ecoregion.Rmd
# Data_08_average_ecoregion_diffs.rds created in LF_08_clim_biome_ecoregion.Rmd

```{r include=FALSE}
library(tidyverse)
library(broom.mixed)
library(kableExtra)
library(ggplot2)
library(sjPlot) ## for the set_theme function
library(viridis)
library(ggpubr)
library(ggeffects)
library(gtools)
library(ggridges)
library(papeR)
library(kableExtra)
library(data.table)
library(cowplot)
library(ggdark)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE,
	fig.height = 8, fig.width = 12)
```

```{r plotting themes}
# set graphical parameters

# blank theme
set_theme(
  base = theme_classic(),
  axis.title.size = 1,
  axis.textsize = 1,
  legend.size = 1,
  legend.title.size = 1,
  geom.label.size = 1,
  plot.margin = margin(1,1,1,1, "pt")
)

## text editing
get_wraper <- function(width) {
  function(x) {
    lapply(strwrap(x, width = width, simplify = FALSE), paste, collapse="\n")}}
## colour palette
cb_pal <- c("#01665e", "#5ab4ac","#c7eae5", "#d8b365", 
            "#8c510a")
```


```{r}
# read in oc model
f_oc_sum <- readRDS("f_oc_sum.rds")
f_oc_sum$land_av <- NA
f_oc_sum$land_av[which(is.na(f_oc_sum$raunk_lf) & !is.na(f_oc_sum$landuse))] <- 1

f_oc_sum$landuse <-gsub(" vegetation", "", f_oc_sum$landuse, ignore.case = TRUE)
f_oc_sum$landuse <- as.character(f_oc_sum$landuse)
f_oc_sum$landuse[f_oc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_oc_sum$landuse <- factor(f_oc_sum$landuse)

## Yvonne's ordering
f_oc_sum$landuse <- factor(f_oc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary",  "Mature secondary", 
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))

## create column for grey bars
f_oc_sum$cat <- as.numeric(f_oc_sum$landuse)
f_oc_sum$rep[f_oc_sum$landuse %in% c("Primary forest",  "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_oc_sum$rep[f_oc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"
```

```{r, include=FALSE}
# read in pc model
f_pc_sum <- readRDS("f_pc_sum.rds")
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)
f_pc_sum$landuse[f_pc_sum$clim == "species richness"] <- "Mature secondary"
f_pc_sum <- f_pc_sum[order(-f_pc_sum$raunk_lf, f_pc_sum$landuse),]
f_pc_sum$land_av <- 0
f_pc_sum$land_av[which(is.na(f_pc_sum$raunk_lf) & !is.na(f_pc_sum$landuse))] <- 1

f_pc_sum$landuse <-gsub(" vegetation", "", f_pc_sum$landuse, ignore.case = TRUE)
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)
f_pc_sum$landuse[f_pc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_pc_sum$landuse <- factor(f_pc_sum$landuse)

f_pc_sum$landuse <- factor(f_pc_sum$landuse)
## Yvonne's ordering
f_pc_sum$landuse <- factor(f_pc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary", "Mature secondary",
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))
f_pc_sum$cat <- as.numeric(f_pc_sum$landuse)
f_pc_sum$rep[f_pc_sum$landuse %in% c("Primary forest", "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_pc_sum$rep[f_pc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"

```

```{r}
## read in model data
if(!exists("ModelDF")) {
  if(file.exists("Data_ModelDF.rds")) {
    try(ModelDF <- readRDS("Data_ModelDF.rds"))
  } else warning("ModelDF does not exist in this directory")
}

ModelDF$Predominant_habitat <-gsub(" vegetation", "", ModelDF$Predominant_habitat, ignore.case = TRUE)
ModelDF$Predominant_habitat <- as.character(ModelDF$Predominant_habitat)
ModelDF$Predominant_habitat[ModelDF$Predominant_habitat == "Secondary (indeterminate age)"] <- "unknown Secondary"
ModelDF$Predominant_habitat <- factor(ModelDF$Predominant_habitat)

## Yvonne's ordering
ModelDF$Predominant_habitat <- factor(ModelDF$Predominant_habitat, 
                      levels = c("Primary forest", "unknown Secondary",  "Mature secondary", 
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))
## get percent cover data for sample sizes table
mydata <- ModelDF

## handle model dataframe to get just percent cover data, with species levels in the right format
levels(ModelDF$Best_guess_binomial) <- gsub(" ", "_", levels(ModelDF$Best_guess_binomial))
mydata <- ModelDF[ModelDF$Diversity_metric == "percent cover", ]
mydata <- mydata[mydata$Measurement !=0,]
mydata <- unique(mydata)
```

*Supplementary Figures for “Land use and climate affect plant occurrence and abundance across life forms”*

## Map of data origins 

```{r, fig.height = 16, fig.width = 24}
source("plot_enviro_pca.R")
biome_rast <- raster("Data_08_anthromes_2_base_data_GeoTIFF/potveg.tif")
biome_rast <- calc(biome_rast, fun=function(x){ x[x <= 0] <- NA; return(x)} )
  par(bg= "black", bty = "l", col="white", col.main = "white", col.lab = "white", col.axis = "white")
plot(biome_rast, col=viridis_pal(option="C")(900))
  points(full_clim[, 'Longitude'], full_clim[,'Latitude'], pch = 21, col = "green", cex = 2)
```

```{r fig.height = 8, fig.width = 10}
biome <- readRDS("Data_08_average_biome_clim_vars.rds")
biome$biome <- factor(biome$biome, 
                      levels = c("Tropical Evergreen Woodland", "Tropical Deciduous Woodland", 
                                 "Dense Shrubland", "Savanna", "Open Shrubland", "Deserts and Barren", 
                                 "Temperate Evergreen Woodland", "Temperate Deciduous Woodland", 
                                 "Grassland and Steppe", "Mixed Woodland", "Boreal Woodland", 
                                 "Tundra"))
  biome_ave <-ggplot(biome, aes(mat, map, colour = biome)) + 
  geom_boxplot(size = 4) +
  labs(colour = "Biome", x = "MAT (degrees C)", y = "MAP (mm)") + 
  scale_color_viridis(discrete = TRUE, option = "plasma") + 
    dark_theme_gray() + theme(legend.position="bottom") 
biome_ave
```


## Appendix S3 – Visualisation of data

```{r, fig.height= 8, fig.width= 12}
## plot model data
lu <- ggplot(ModelDF, aes(x=Predominant_habitat)) +
    geom_bar(aes(x=Predominant_habitat),  fill = "white", size=1, alpha = 1, show.legend = F) +
  geom_bar(data = ModelDF[ModelDF$Diversity_metric == "percent cover" & ModelDF$Measurement != 0,], 
               aes(x=Predominant_habitat),  size=1, alpha = 1, show.legend = F) +
       theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
               scale_x_discrete(labels = get_wraper(10)) +
  labs(title = "A", x = "Land use") 

raunk <- ggplot(ModelDF, aes(x=raunk_lf)) +
    geom_bar(aes(x=raunk_lf),  fill = "white", size=1, alpha = 1, show.legend = F) +
  geom_bar(data = ModelDF[ModelDF$Diversity_metric == "percent cover" & ModelDF$Measurement != 0,], 
               aes(x=raunk_lf),  size=1, alpha = 1, show.legend = F) +
       theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
               scale_x_discrete(labels = get_wraper(10)) +
  labs(title = "B", x = "Life form") 

lu2 <- ggplot(ModelDF[ModelDF$Diversity_metric == "percent cover" & ModelDF$Measurement != 0,],
              aes(x=Predominant_habitat)) +
    geom_bar(aes(x=Predominant_habitat), size=1, alpha = 1, show.legend = F) +
       theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
               scale_x_discrete(labels = get_wraper(10),drop=FALSE) +
  labs(title = "C", x = "Land use") + ylim(0, 12500)
       
raunk2 <- ggplot(ModelDF[ModelDF$Diversity_metric == "percent cover" & ModelDF$Measurement != 0,], aes(x=raunk_lf)) +
    geom_bar(aes(x=raunk_lf),  size=1, alpha = 1, show.legend = F) +
       theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
               scale_x_discrete(labels = get_wraper(10)) +
  labs(title = "D", x = "Life form") + ylim(0, 12500)
```

```{r, fig.height= 8, fig.width= 12}
lu <- lu + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
                 axis.text.x = element_blank(),  axis.ticks.x = element_blank())
lu2 <- lu2 + theme(axis.title.y = element_blank()) 
raunk <- raunk + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                         axis.ticks.x = element_blank(), axis.title.y = element_blank())
raunk2 <- raunk2 + theme(axis.title.y = element_blank()) 

resp <- ggarrange(lu, raunk, lu2, raunk2,  nrow = 2, ncol = 2, align = "hv", heights = c(2, 1)) + theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))
annotate_figure(resp, #top = text_grob("Relative proportions of model data", face = "bold", size = 20),
                left = text_grob("Count", size = 15, rot = 90))
```

```{r eval=FALSE, include=FALSE}
sample_sizes <- as.data.frame(t(c(1:3)))
for (i in levels(ModelDF$Predominant_habitat)){
  for (j in levels(ModelDF$raunk_lf)){
  sample_sizes <- rbind(sample_sizes, c(i,j, length(ModelDF$Predominant_habitat[ModelDF$Predominant_habitat == i & ModelDF$raunk_lf == j])))
  }}
names(sample_sizes) <- c("Land use", "Life form", "Occurrence sample size")

for (i in levels(ModelDF$Predominant_habitat)){
  for (j in levels(ModelDF$raunk_lf)){
sample_sizes$'Abundance sample size'[sample_sizes$`Land use` == i & sample_sizes$`Life form` == j] <-
  length(mydata$Predominant_habitat[mydata$Predominant_habitat == i & mydata$raunk_lf == j])
}}

sum.df <- as_tibble(sample_sizes[-1,])
sum.df %>%
  kable(booktabs = T, caption = "Table 10. Model sample sizes") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  kable_styling(font_size = 10)
```

```{r eval=FALSE, include=FALSE}
sample_sizes <- as.data.frame(t(c(1:3)))
for (i in levels(mydata$Predominant_habitat)){
  for (j in levels(mydata$raunk_lf)){
  sample_sizes <- rbind(sample_sizes, c(i,j, length(mydata$Predominant_habitat[mydata$Predominant_habitat == i & mydata$raunk_lf == j])))
  }}
names(sample_sizes) <- c("Land use", "Life form", "Sample size")

sum.df <- as_tibble(sample_sizes[-1,])
sum.df %>%
  kable(booktabs = T, caption = "Table 10b. Abundance model sample sizes") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  kable_styling(font_size = 10)
```

## Data validation extended methods 

## Biome and Ecoregion visual summaries
```{r fig.height = 4, fig.width = 10}
biome <- readRDS("Data_08_average_biome_clim_vars.rds")

## make plot
biome_ave <-ggplot(biome, aes(mat, map, colour = biome)) + 
  geom_point(size = 4) +
  geom_pointrange(aes(ymin=map - map_var, ymax=map + map_var)) + 
  labs(colour = "Biome", x = "MAT (degrees C)", y = "MAP (mm)") + 
  scale_color_viridis(discrete = TRUE)

biome_var_ave <-ggplot(biome, aes(mat, mat_var, colour = biome)) + 
  geom_point(size = 4) +
  labs(colour = "Biome", x = "MAT (degrees C)", y = "MAT_var (degrees C)") + 
  scale_color_viridis(discrete = TRUE)

resp <-ggarrange(biome_ave, biome_var_ave, nrow = 1, ncol = 2, align = "hv", heights = c(1, 1), common.legend = TRUE, legend = "right") +
  theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm")) 
annotate_figure(resp, top = text_grob("Biome averages for climatic variables", face = "bold", size = 20))

```

```{r fig.height = 4, fig.width = 6}
eco <- readRDS("Data_08_average_ecoregion_clim_vars.rds")

## make plot
eco_ave <-ggplot(eco, aes(mat, map, colour = mat_var)) + 
  geom_point(size = 4) +
  geom_pointrange(aes(ymin=map - map_var, ymax=map + map_var)) + 
  labs(title = "Ecoregion averages for climatic variables", colour = "MAT_var", x = "MAT (degrees C)", y = "MAP (mm)") + 
  scale_color_viridis()


resp <-ggarrange(eco_ave, nrow = 1, ncol = 1) +
  theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm")) 
annotate_figure(resp, top = text_grob("Ecoregion averages for climatic variables", face = "bold", size = 20))

```

```{r fig.height = 6.5, fig.width = 9}
data <- readRDS("Data_08_sd_model_clim_data.rds")
biome_diff <- readRDS("Data_08_average_biome_diffs.rds")
eco_diff <- readRDS("Data_08_average_ecoregion_diffs.rds")
vars <- c("map", "mat", "map_var", "mat_var")

## plot comparison of average differences with 1 sd in model data
par(mfrow =c(3,4), mar=c(3,2,3,2))

for(i in names(data[,which(names(data) %in% vars)])){
  par(bty = "l")
  hist(data[,i], 
       breaks = 300,
       border = "grey", ylab = NULL, xlab = NULL,main = NULL, freq = F)
  lines(density(data[,i]))
  abline(v=mean(data[,i]),col="black", lwd = 2)
  abline(v=median(data[,i]),lty=2,col="black", lwd = 2)
  abline(v = mean(data[,i])+sd(data[,i]), lty = 1, col = "#A715AD", lwd = 2)
  abline(v = mean(data[,i])-sd(data[,i]), lty = 1, col = "#A715AD", lwd = 2)}

for(i in names(biome_diff[,which(names(biome_diff) %in% vars)])){
  par(bty = "l")
  hist(biome_diff[,i],
       breaks = 300, 
       border = "grey",ylab = NULL, xlab = NULL, main = NULL, freq = F)
  lines(density(biome_diff[,i]))
    abline(v=mean(biome_diff[,i]),col="black", lwd = 2)
  abline(v=median(biome_diff[,i]),lty=2,col="black", lwd = 2)
  abline(v = sd(data[,i]), lty = 1, col = "#A715AD", lwd = 2)}

for(i in names(eco_diff[,which(names(eco_diff) %in% vars)])){
  par(bty = "l")
  hist(eco_diff[,i],
       breaks = 300,
       border = "grey", ylab = NULL,main = NULL, xlab = paste(i), freq = F)
  lines(density(eco_diff[,i]))
    abline(v=mean(eco_diff[,i]),col="black", lwd = 2)
  abline(v=median(eco_diff[,i]),lty=2,col="black", lwd = 2)
  abline(v = sd(data[,i]), lty = 1, col = "#A715AD", lwd = 2)}

```










