---
title: "LF_11_sup_tables"
author: "Caroline McKeon"
date: "10/13/2021"
output:
  html_document:
    df_print: paged
---


## DATA NEEDED:

# f_oc_sum - dataframe of results from occurrence model object - created in LF_06b_oc_estimates.R
# f_pc_sum - dataframe of results from abundance model object - created in LF_06a_pc_estimates.R

# Data_ModelDF.rds  - Model dataframe with 624696 obs of 24 variables, created in LF_01_data_handling.R

# Data_08_all_clim_ave.rds created in LF_08_clim_biome_ecoregion.Rmd
# Data_08_average_biome_clim_vars.rds created in LF_08_clim_biome_ecoregion.Rmd

# f_oc_NO_RICH_sum - dataframe of results from NO RICHNESS occurrence model object - created in LF_07b_oc_effectsizes.R
# f_pc_NO_RICH_sum - dataframe of results from NO RICHNESS abundance model object - created in LF_07a_pc_effectsizes.R


```{r echo=FALSE}
library(tidyverse)
library(broom.mixed)
library(data.table)
library(ggplot2)
library(sjPlot) ## for the set_theme function
library(viridis)
library(ggpubr)
library(ggeffects)
library(gtools)
library(ggridges)
library(papeR)
library(kableExtra)
library(cowplot)
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	fig.height = 9.7,
	fig.width = 16
)
knitr::opts_chunk$set(fig.width=16, fig.height=9.7) 
```

## read in data
```{r}
# read in oc model
f_oc_sum <- readRDS("f_oc_sum.rds")
f_oc_sum$land_av <- NA
f_oc_sum$land_av[which(is.na(f_oc_sum$raunk_lf) & !is.na(f_oc_sum$landuse))] <- 1

f_oc_sum$landuse <-gsub(" vegetation", "", f_oc_sum$landuse, ignore.case = TRUE)
f_oc_sum$landuse <- as.character(f_oc_sum$landuse)
f_oc_sum$landuse[f_oc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_oc_sum$landuse <- factor(f_oc_sum$landuse)

## Yvonne's ordering
f_oc_sum$landuse <- factor(f_oc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary",  "Mature secondary", 
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))

## create column for grey bars
f_oc_sum$cat <- as.numeric(f_oc_sum$landuse)
f_oc_sum$rep[f_oc_sum$landuse %in% c("Primary forest",  "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_oc_sum$rep[f_oc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"
```

```{r, include=FALSE}
# read in pc model
f_pc_sum <- readRDS("f_pc_sum.rds")
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)
f_pc_sum$landuse[f_pc_sum$clim == "species richness"] <- "Mature secondary"
f_pc_sum <- f_pc_sum[order(-f_pc_sum$raunk_lf, f_pc_sum$landuse),]
f_pc_sum$land_av <- 0
f_pc_sum$land_av[which(is.na(f_pc_sum$raunk_lf) & !is.na(f_pc_sum$landuse))] <- 1

f_pc_sum$landuse <-gsub(" vegetation", "", f_pc_sum$landuse, ignore.case = TRUE)
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)
f_pc_sum$landuse[f_pc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_pc_sum$landuse <- factor(f_pc_sum$landuse)

f_pc_sum$landuse <- factor(f_pc_sum$landuse)
## Yvonne's ordering
f_pc_sum$landuse <- factor(f_pc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary", "Mature secondary",
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))
f_pc_sum$cat <- as.numeric(f_pc_sum$landuse)
f_pc_sum$rep[f_pc_sum$landuse %in% c("Primary forest", "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_pc_sum$rep[f_pc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"

```

```{r}
## read in model data
if(!exists("ModelDF")) {
  if(file.exists("Data_ModelDF.rds")) {
    try(ModelDF <- readRDS("Data_ModelDF.rds"))
  } else warning("ModelDF does not exist in this directory")
}

ModelDF$Predominant_habitat <-gsub(" vegetation", "", ModelDF$Predominant_habitat, ignore.case = TRUE)
ModelDF$Predominant_habitat <- as.character(ModelDF$Predominant_habitat)
ModelDF$Predominant_habitat[ModelDF$Predominant_habitat == "Secondary (indeterminate age)"] <- "unknown Secondary"
ModelDF$Predominant_habitat <- factor(ModelDF$Predominant_habitat)

## Yvonne's ordering
ModelDF$Predominant_habitat <- factor(ModelDF$Predominant_habitat, 
                      levels = c("Primary forest", "unknown Secondary",  "Mature secondary", 
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))
```

```{r}
f_oc_sum$'Model terms' <- paste(f_oc_sum$landuse, ":", f_oc_sum$raunk_lf, ":", f_oc_sum$clim)
nope <- c("\\: 0", " \\: 0", "NA \\: ")
f_oc_sum$'Model terms' <- gsub(paste(nope, collapse="|"), "", f_oc_sum$'Model terms')
f_oc_sum$'Model terms' <- gsub(" \\: 0", "", f_oc_sum$'Model terms')
f_oc_sum$'Model terms'[f_oc_sum$`Model terms` == "0"] <- "Grand Mean"

f_pc_sum$'Model terms' <- paste(f_pc_sum$landuse, ":", f_pc_sum$raunk_lf, ":", f_pc_sum$clim)
nope <- c("\\: 0", " \\: 0", "NA \\: ")
f_pc_sum$'Model terms' <- gsub(paste(nope, collapse="|"), "", f_pc_sum$'Model terms')
f_pc_sum$'Model terms' <- gsub(" \\: 0", "", f_pc_sum$'Model terms')
f_pc_sum$'Model terms'[f_pc_sum$`Model terms` == "0"] <- "Grand Mean"

```


# Table 3. Occurrence model output
```{r}
# tables of model outputs
sum.df <- as_tibble(f_oc_sum[,c(15,2:5)])

sum.df %>% 
  kable(booktabs = T, caption = "Table 1. Occurrence model output") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  row_spec(which(sum.df$`z value` < -2), color = "red", background = "white") %>%
  row_spec(which(sum.df$`z value` >2), color = "blue", background = "white") %>% 
  kable_styling(font_size = 10)

```

# Table 4. Abundance model output
```{r}
sum.df <- as_tibble(f_pc_sum[,c(16,2:4)])

sum.df %>% 
  kable(booktabs = T, caption = "Table 2. Abundance model outputt") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  row_spec(which(sum.df$`t value` < -2), color = "red", background = "white") %>%
  row_spec(which(sum.df$`t value` >2), color = "blue", background = "white") %>% 
  kable_styling(font_size = 10)

```

# Table 5a. Effect size for Life form Occurrence by Land use
```{r}
## effect size tables Occurrence
oc_eff_lu <- f_oc_sum[which(!is.na(f_oc_sum$landuse) & !is.na(f_oc_sum$raunk_lf))]
oc_eff_lu$Magnitude <- oc_eff_lu$eff_sz - f_oc_sum$Estimate[f_oc_sum$rn == "(Intercept)"]

oc_eff_lu  <- oc_eff_lu[,c(9:10,16)]
oc_eff_lu <- oc_eff_lu[order(oc_eff_lu$Magnitude),]

sum.df <- as_tibble(oc_eff_lu)
sum.df %>% 
  kable(booktabs = T, caption = "Table 3a. Effect size for Life form Occurrence by Land use") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```

# Table 5b. Effect size for Life form Occurrence by Climate
```{r}
oc_eff_clim <- f_oc_sum[which(!is.na(f_oc_sum$raunk_lf) & f_oc_sum$clim != 0 & f_oc_sum$clim != "species richness")]
oc_eff_clim$Magnitude <- oc_eff_clim$eff_sz

oc_eff_clim  <- oc_eff_clim[,c(10:11,16)]
oc_eff_clim <- oc_eff_clim[order(oc_eff_clim$Magnitude),]

sum.df <- as_tibble(oc_eff_clim)
sum.df %>% 
  kable(booktabs = T, caption = "Table 3b. Effect size for Life form Occurrence by Climate") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```

## Table 5c. Range within effect sizes
```{r}
# d <- f_oc_sum[which(!is.na(f_oc_sum$landuse) & !is.na(f_oc_sum$raunk_lf)),]
# d$landuse <- factor(d$landuse)
# for(i in levels(d$landuse)){
#   d$land_range[d$landuse == i] <- max(d$eff_sz[d$landuse == i]) - min(d$eff_sz[d$landuse == i])
# }
# 
# d <- unique(d[, c(9,16)])
# d <- d[order(d$land_range),]
# sum.df <- as_tibble(d)
# sum.df %>% 
#   kable(booktabs = T, caption = "Table 3c. Range in Effect size for Life form Occurrence by Land use") %>%
#   kable_styling(bootstrap_options = "striped", full_width = T) 
# 
# c <- f_oc_sum[which(!is.na(f_oc_sum$raunk_lf) & f_oc_sum$clim != 0 & f_oc_sum$clim != "species richness")]
# c$clim<- factor(c$clim)
# for(i in levels(c$clim)){
#   c$clim_range[c$clim == i] <- max(c$eff_sz[c$clim == i]) - min(c$eff_sz[c$clim == i])
# }
# 
# c <- unique(c[, c(11,16)])
# c <- c[order(c$clim_range),]
# sum.df <- as_tibble(c)
# sum.df %>% 
#   kable(booktabs = T, caption = "Table 3d. Range in Effect sizes for Life form Occurrence by Climate") %>%
#   kable_styling(bootstrap_options = "striped", full_width = T) 

g <- f_oc_sum[which(!is.na(f_oc_sum$landuse) & !is.na(f_oc_sum$raunk_lf)),]
for(i in levels(g$raunk_lf)){
  g$raunk_lf_range[g$raunk_lf == i] <- max(g$eff_sz[g$raunk_lf == i]) - min(g$eff_sz[g$raunk_lf == i])
}
g <- unique(g[, c(10,16)])
g <- g[order(g$raunk_lf_range),]
sum.df <- as_tibble(g)
sum.df %>% 
  kable(booktabs = T, caption = "Table 3e. Range in Effect sizes for Life form Occurrence across Land use") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```


# Table 6a. Magnitude of Effect size for Life form Abundance by Land use
```{r}
## effect size tables Abundance
pc_eff_lu <- f_pc_sum[which(!is.na(f_pc_sum$landuse) & !is.na(f_pc_sum$raunk_lf))]
pc_eff_lu$Magnitude <- pc_eff_lu$eff_sz - f_pc_sum$Estimate[f_pc_sum$rn == "(Intercept)"]

pc_eff_lu  <- pc_eff_lu[,c(8:9,17)]
pc_eff_lu <- pc_eff_lu[order(pc_eff_lu$Magnitude),]

sum.df <- as_tibble(pc_eff_lu)
sum.df %>% 
  kable(booktabs = T, caption = "Table 4a. Effect size for Life form Abundance by Land use") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```

# Table 6b. Magnitude of Effect size for Life form Abundance by Climate
```{r}
pc_eff_clim <- f_pc_sum[which(!is.na(f_pc_sum$raunk_lf) & f_pc_sum$clim != 0 & f_pc_sum$clim != "species richness")]
pc_eff_clim$Magnitude <- pc_eff_clim$eff_sz

pc_eff_clim  <- pc_eff_clim[,c(9:10,17)]
pc_eff_clim <- pc_eff_clim[order(pc_eff_clim$Magnitude),]

sum.df <- as_tibble(pc_eff_clim)
sum.df %>% 
  kable(booktabs = T, caption = "Table 4b. Effect size for Life form Abundance by Climate") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```

## Table 6c. Range within effect sizes
```{r}
# d <- f_pc_sum[which(!is.na(f_pc_sum$landuse) & !is.na(f_pc_sum$raunk_lf)& f_pc_sum$clim != "species richness"),]
# d$landuse <- factor(d$landuse)
# for(i in levels(d$landuse)){
#   d$land_range[d$landuse == i] <- max(d$eff_sz[d$landuse == i]) - min(d$eff_sz[d$landuse == i])
# }
# d <- unique(d[, c(8,17)])
# d <- d[order(d$land_range),]
# 
# sum.df <- as_tibble(d)
# sum.df %>% 
#   kable(booktabs = T, caption = "Table 4c. Range in Effect size for Life form Abundance by Land use") %>%
#   kable_styling(bootstrap_options = "striped", full_width = T) 
# 
# c <- f_pc_sum[which(!is.na(f_pc_sum$raunk_lf) & f_pc_sum$clim != 0 & f_pc_sum$clim != "species richness")]
# c$clim<- factor(c$clim)
# for(i in levels(c$clim)){
#   c$clim_range[c$clim == i] <- max(c$eff_sz[c$clim == i]) - min(c$eff_sz[c$clim == i])
# }
# c <- unique(c[, c(10,17)])
# c <- c[order(c$clim_range),]
# 
# sum.df <- as_tibble(c)
# sum.df %>% 
#   kable(booktabs = T, caption = "Table 4d. Range in Effect sizes for Life form Abundance by Climate") %>%
#   kable_styling(bootstrap_options = "striped", full_width = T) 

g <- f_pc_sum[which(!is.na(f_pc_sum$landuse) & !is.na(f_pc_sum$raunk_lf)& f_pc_sum$clim != "species richness"),]
for(i in levels(g$raunk_lf)){
  g$raunk_lf_range[g$raunk_lf == i] <- max(g$eff_sz[g$raunk_lf == i]) - min(g$eff_sz[g$raunk_lf == i])
}
g <- unique(g[, c(9,17)])
g <- g[order(g$raunk_lf_range),]

sum.df <- as_tibble(g)
sum.df %>% 
  kable(booktabs = T, caption = "Table 4e. Range in Effect sizes for Life form Abundance across Land use") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) 
```

# Table 7. Summary statistics for climatic variables
```{r}
all_clim_ave <- readRDS("Data_08_all_clim_ave.rds")
opts <- options(knitr.kable.NA = "")
sum.df <- as_tibble(all_clim_ave)

sum.df %>% 
  kable(booktabs = T, caption = "Table 5. Summary statistics for climatic variables") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) 
```

# Table 8. Biome averages for climatic variables
```{r}
biome <- readRDS("Data_08_average_biome_clim_vars.rds")

sum.df <- as_tibble(biome[,c(5,1:4)])

sum.df %>% 
  kable(booktabs = T, caption = "Table 6. Biome averages for climatic variables") %>%
  kable_styling(bootstrap_options = "striped", full_width = F) 
```


```{r}
# read in oc model
f_oc_sum <- readRDS("f_oc_NO_RICH_sum.rds")
f_oc_sum$land_av <- NA
f_oc_sum$land_av[which(is.na(f_oc_sum$raunk_lf) & !is.na(f_oc_sum$landuse))] <- 1

f_oc_sum$landuse <-gsub(" vegetation", "", f_oc_sum$landuse, ignore.case = TRUE)
f_oc_sum$landuse <- as.character(f_oc_sum$landuse)
f_oc_sum$landuse[f_oc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_oc_sum$landuse <- factor(f_oc_sum$landuse)

## Yvonne's ordering
f_oc_sum$landuse <- factor(f_oc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary",  "Mature secondary", 
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))

## create column for grey bars
f_oc_sum$cat <- as.numeric(f_oc_sum$landuse)
f_oc_sum$rep[f_oc_sum$landuse %in% c("Primary forest",  "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_oc_sum$rep[f_oc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"
```

```{r, include=FALSE}
# read in pc model
f_pc_sum <- readRDS("f_pc_NO_RICH_sum.rds")
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)

f_pc_sum <- f_pc_sum[order(-f_pc_sum$raunk_lf, f_pc_sum$landuse),]
f_pc_sum$land_av <- 0
f_pc_sum$land_av[which(is.na(f_pc_sum$raunk_lf) & !is.na(f_pc_sum$landuse))] <- 1

f_pc_sum$landuse <-gsub(" vegetation", "", f_pc_sum$landuse, ignore.case = TRUE)
f_pc_sum$landuse <- as.character(f_pc_sum$landuse)
f_pc_sum$landuse[f_pc_sum$landuse == "Secondary (indeterminate age)"] <- "unknown Secondary"
f_pc_sum$landuse <- factor(f_pc_sum$landuse)

f_pc_sum$landuse <- factor(f_pc_sum$landuse)
## Yvonne's ordering
f_pc_sum$landuse <- factor(f_pc_sum$landuse, 
                      levels = c("Primary forest", "unknown Secondary", "Mature secondary",
                                 "Intermediate secondary", "Young secondary",
                                 "Plantation forest", "Primary non-forest", "Pasture", "Cropland", "Urban"))
f_pc_sum$cat <- as.numeric(f_pc_sum$landuse)
f_pc_sum$rep[f_pc_sum$landuse %in% c("Primary forest", "Mature secondary", "Young secondary",
                                     "Primary non-forest", "Cropland")] <- "white"
f_pc_sum$rep[f_pc_sum$landuse %in% c("unknown Secondary","Intermediate secondary",
                                     "Plantation forest", "Pasture", "Urban")] <- "grey"

```

```{r}
f_oc_sum$'Model terms' <- paste(f_oc_sum$landuse, ":", f_oc_sum$raunk_lf, ":", f_oc_sum$clim)
nope <- c("\\: 0", " \\: 0", "NA \\: ")
f_oc_sum$'Model terms' <- gsub(paste(nope, collapse="|"), "", f_oc_sum$'Model terms')
f_oc_sum$'Model terms' <- gsub(" \\: 0", "", f_oc_sum$'Model terms')
f_oc_sum$'Model terms'[f_oc_sum$`Model terms` == "0"] <- "Grand Mean"

f_pc_sum$'Model terms' <- paste(f_pc_sum$landuse, ":", f_pc_sum$raunk_lf, ":", f_pc_sum$clim)
nope <- c("\\: 0", " \\: 0", "NA \\: ")
f_pc_sum$'Model terms' <- gsub(paste(nope, collapse="|"), "", f_pc_sum$'Model terms')
f_pc_sum$'Model terms' <- gsub(" \\: 0", "", f_pc_sum$'Model terms')
f_pc_sum$'Model terms'[f_pc_sum$`Model terms` == "0"] <- "Grand Mean"

```


# Table 9a. Occurrence no species richness model output
```{r}
# tables of model outputs
sum.df <- as_tibble(f_oc_sum[,c(16,2:5)])

sum.df %>% 
  kable(booktabs = T, caption = "Table 1. Occurrence model output") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  row_spec(which(sum.df$`z value` < -2), color = "red", background = "white") %>%
  row_spec(which(sum.df$`z value` >2), color = "blue", background = "white") %>% 
  kable_styling(font_size = 10)

```

# Table 9b. Abundance no species richness model output
```{r}
sum.df <- as_tibble(f_pc_sum[,c(16,2:4)])

sum.df %>% 
  kable(booktabs = T, caption = "Table 2. Abundance model outputt") %>%
  kable_styling(bootstrap_options = "striped", full_width = T) %>%
  row_spec(which(sum.df$`t value` < -2), color = "red", background = "white") %>%
  row_spec(which(sum.df$`t value` >2), color = "blue", background = "white") %>% 
  kable_styling(font_size = 10)

```



And finally, get all R package citations.
```{r}
list <- c("ggdark", "broom.mixed", "cowplot", "gdalUtils", "kableExtra", 
          "papeR", "xtable", "car", "carData", "ggridges", "gtools", "ggeffects", 
          "ggpubr", "viridis", "viridisLite", "sjPlot", "data.table", "DHARMa", 
          "wec", "glmmTMB", "optimx", "lme4", "Matrix", "RColorBrewer", 
          "rgdal", "raster", "sp", "forcats", "stringr", "dplyr", "purrr", 
          "readr", "tidyr", "tibble", "ggplot2", "tidyverse", "stats", 
          "graphics", "grDevices", "utils", "datasets", "methods", "base")
lapply(list, citation)
```






